---
title: Working with tmux and R on remote computers
author: Xiang Zhang
date: '2021-05-18'
slug: []
categories:
  - R
tags: []
---
This is my first-ever blog post. 
For the purpose of network reverse-engineering, I need to work with my powerful workstation at the university campus. 
Since I am working from home, I need to do that remotely. 
Furthermore, since the programs that I need for network reverse-engineering are Bioconductor packages, I also need to install these packages beforehand. 
I am a disciplined R programmer. 
I always use `renv` to manage packages for each project individually. 
Can I keep my discipline when I work with the remote workstation as well? 
In this post, I will record my steps from scratch. 

# Step 1 Connect to the remote workstation

```{bash eval=FALSE}
ssh xiang@alive.bio.uu.nl
```

Then type the password. 
Next, log in my workstation. 

```{bash eval=FALSE}
ssh trias
```

# Step 2 Start a tmux session

I have heard `tmux` for a long time. 
The most important reason people love `tmux` is that your computation will not be killed when you lose the internet connection. 
Here I start a new `tmux` session.

```{bash eval=FALSE}
tmux
```

# Step 3 Create a project folder

I have a folder called "Projects" in my home directory. 
Now, I am going to make a new folder inside the "Projects" folder to include all the data and scripts I use for the network reverse-engineering. 

```{bash eval=FALSE}
cd Projects
mkdir lateral_root_formation
cd lateral_root_formation
```

# Step 4 Install `renv` package

Now, I am standing in the folder "lateral_root_formation" and I am ready to install Bioconductor packages. 
I always use `renv` to make sure isolation of project-specific packages. 
How can I do that here?

First, I launch the R

```{bash eval=FALSE}
R
```

Next, I will install the `renv` package.

```{r eval=FALSE}
install.packages("renv")
```

Once the installation is done, I am going to initialize the `renv`. 

```{r eval=FALSE}
renv::init()
```

In the message of `renv::init()` I am told to restart the R session. 
So I did that. 

# Step 5 Install Bioconductor packages

Now I am ready to install Bioconductor packages. 
For this project, I am going to use the Bioconductor `GRENITS`. 

```{r eval=FALSE}
install.packages("BiocManager")
BiocManager::install("GRENITS")
```

# Step 6 Run a network inference example on the remote workstation

Now it's time to have a try. 
Here I am going to run the example that is presented in the documentation of `GRENITS` package. 
For working on remote workstation, I need to have a R script and run `Rscript`. 

First, I quit from the R interactive session and I created a R script with the text editor, `nano`. 

```{bash eval=FALSE}
touch run_grenits_example.R
nano run_grenits_example.R
```

The R script contains the following R commands

```{r eval=FALSE}
#! /usr/bin/Rscript
library(GRENITS)

data(Athaliana_ODE)

output.folder <- "./Example_LinearNet"
LinearNet(output.folder, Athaliana_ODE)

# Analyse raw results, place analysis plots and files in output.folder
analyse.output(output.folder)
```

To run the script with Rscript, I must first authorize the file:

```{bash eval=FALSE}
chmod +x run_grenits_example.R
```

Then, I can execute the R commands.

```{bash eval=FALSE}
Rscript run_grenits_example.R
```

For this particular example, the computational time is in seconds, so there is no obvious advantage of using `tmux` in this case. 
However, when the calculation will take much longer time, then what I will do is to press `Ctrl` + `B` and then press `D` to detach the session. 
The magical thing is that the computation will go on while I am driving home from work. 
Once I get home and re-connect the internet, I can check the status of the computation by 

```{bash eval=FALSE}
ssh xiang@alive.bio.uu.nl
ssh trias
tmux ls
tmux attach -t 0 # or the name of the session
```

# Step 7 Transfer the outcome files to my local computer

My laptop is my primary workhorse for data visualization, writing manuscript and making presentations. 
Therefore, I need frequently download the computational outcomes from the workstation to my local computer. 
How to do this?

First, start a new terminal session on my laptop. 
Then go to the directory where I want to keep the output files. 

```{bash eval=FALSE}
cd Downloads/
scp xiang@alive.bio.uu.nl:/home/xiang/Projects/lateral_root_formation/Example_LinearNet/"*" .
```

Type the password and then I get all the files I want!

# The end

This is my first post and I start with a note for my own. 
This is a simple workflow that I probably will do many times later. 
If I forget something or some commands, I will read my own post and won't bother my colleagues again with the same questions. 
If it can help someone, I will be more than happy. 
